---
title: "Hawaii Solar Irradiance"
author: "Math 271"
date: "Spring 2022"
output: 
  html_document:
    css: lab.css
    keep_md: true
    theme: lumen
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Solution document setup

Create a solution document and save it as `solution.Rmd`.

Get in the habit of loading all packages in a setup chunk at the top of the file. Change the chunk options area to have `include=FALSE, warning=FALSE, message=FALSE` to hide all the loading chatter, as shown in this instruction document.

## Daily Solar Irradiance for Hawaii

Here is a published collection of data of [solar irradiance](https://figshare.com/articles/dataset/Daily_Incoming_Solar_Radiation_in_Hawaii/5579134)
for the state of Hawaii.

Below is an example of how to download a file from a URL and save it to whatever computer runs the Rmd file. The `if()` statement ensures that the file is not downloaded repeatedly. This is an example of _conditional code execution_. Copy and paste this code into your document.

```{r import_data}
solar_file <- "daily_solar_radiation.txt"
if( !file.exists(solar_file) )
  download.file("https://ndownloader.figshare.com/files/9700246", solar_file)

raw_solar <- read_csv(solar_file)
```

This file is a typical example of raw data that is not in tidy form. By perusing the raw data file I was able to determine that it has several columns of reasonably tidy information, such as measuring station identification tags, Latitude, Longitude, and Elevation. However, then it has thousands of columns with names like `X1990.05.10`, and these columns contain the irradiance data. There are many `NA` (missing data) values for the irradiance. Presumably, some stations did not exist for the entire period of time.

```{r}
#str(raw_solar, list.len=15, give.attr=FALSE)
raw_solar
```

## Pivot and wrangle

1. Use `tidyr::pivot_longer` to pivot all columns that start with "X" into new variables `date` and `irradiance`, dropping all the `NA` values. Save the result in `solar`. (Read the `?pivot_longer` help file.)

2. Fix column names. Some of the column names are a bit ugly and annoying to type. Use `dplr::rename` to rename these columns
    - `Station.Name` to `name`
    - `Sta..ID` to `id`
    - `LON` to `lon`
    - `LAT` to `lat`
    - `Elev.` to `elev`

Note the consistency in capitalization. This is worth your time.

## Time is an illusion, lunchtime doubly so

The remaining portion of the new `date` variable looks like `1990.05.10`, and is a character string. But this string obviously represents a point in time in Year.Month.Day format. _The._ __*Horror.*__

[__Time is a data nightmare.__](https://www.mojotech.com/blog/the-complexity-of-time-data-programming/)
That link only gives you the vague outlines of just the issues that humans have caused through political fractiousness and our insistence on syncing clocks with the passage of the sun in the sky. 
[Nature](https://en.wikipedia.org/wiki/Solar_time) has even 
[more fun](https://en.wikipedia.org/wiki/Sidereal_time) 
and [surprises](https://en.wikipedia.org/wiki/Terrestrial_Time)
in [store for us](https://en.wikipedia.org/wiki/Introduction_to_general_relativity), although those issues are of less practical importance. The primary source of difficulties for most Earthlings are time zones, daylight savings time, and leap days (and leap _seconds_ if you're really into precision). 

By the way, are all your clocks synchronized?

The tidyverse tool for handling dates and times is the `lubridate` package. 
It lubricates dates. Have a gaze at the
[Lubridate cheat sheet](https://rawgit.com/rstudio/cheatsheets/master/lubridate.pdf). Spend some time being boggled by the section _Math with Date-times_ on the second page. If you thought you knew what _11/4/2018 01:30 US/Eastern_ __really__ means, or if you thought that _six hours_ was a pretty cut and dry concept, then I hope you are feeling suitably chastened.


3. Parse the `Date` strings. Read the help and then use the `lubridate::ymd` function with `mutate` to replace the `date` column with interpreted R date objects. 

4. Create a new `Month` variable Read the help and then use `lubridate::month` with `mutate` to create a new variable with the `month` of the observation. Use the appropriate argument to `month` to get names instead of numbers.

:::{.license}
Make sure that you've stored the results of your wrangling manipulations back to the `solar` object. You can either do this by making one big pipe `pivot_longer(...) %>% select(...) %>% mutate(...) %>% ...` or by doing it one line at a time, updating the `solar` object as you go.
:::

## Data exploration

The next couple of exercises are answered with code. You can use a chunk option `result="hide"` to supress the output in your document. This is good for demonstrating code without showing unneeded output.

5. Practice by creating `filter` views of the data on these conditions:
    - Big Island stations only
    - Neighbor island stations (not Oahu)
    - Stations below 1000 ft elevation
    - Stations between 1000 and 2000 ft elevation
    - Measurements in June, July, or August.
    - Big Island stations with a Longitude west of -155.55 degrees. (Basically, West Hawaii.)
 
6. All-data summaries. Use `summarize` to find:
    - `mean` irradiance and its `sd` (standard deviation)
    - `min` and `max` station elevation
 
7. Group summaries. Use `group_by` with `tally` or `summarize` to find:
    - mean irradiance on each island.
    - mean irradiance each month.
    - number of observations for each station id.
    - number of observations for each combination of Island and Network.
    - The date of the first and last record at each station. 
      + Cue Jaws theme music.
        + It's not really that bad. Use `min` and `max`.

Answer this one with code and text.

8. Here's one more question that you can answer fairly easily with these methods, although I'll leave it up to you to try and work out a way to do it: _Do the Station Names and Station IDs contain completely redundant information?_ Or stated another way: _Is Station ID just a code for the Station Name?_

## Submitting your answers

- Make sure you pass all the autograding tests in the `tests.R` file.
- Make sure that you have committed the solution `.Rmd` file to git.
- Push everything to github. Double-check things there. If you used any extra packages make sure you added them to the PACKAGES file so that the autograding script can install them.